---
title: "Homework 6"
author: "Christopher Crowe"
date: "November 16, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

set.seed(1)

```

### Problem 1

The code chunk below puts the data in a format that is appropriate for modeling. For example, it ensures that predictors of interest are the appropriate data types and removes data from some cities that do not report some data.

```{r}
homicides = read.csv("./data/homicide-data.csv")

homicides = 
  homicides %>% 
  mutate(city_state = paste(city, state, sep = ", "),
         resolved = as.numeric(disposition == "Closed by arrest"),
         victim_race_2 = case_when(
           victim_race == "Asian" ~ "Non-White",
           victim_race == "Black" ~ "Non-White",
           victim_race == "Hispanic" ~ "Non-White",
           victim_race == "Other" ~ "Non-White",
           victim_race == "Unknown" ~ "Non-White",
           victim_race == "White" ~ "White"
         ),
         victim_race = factor(victim_race_2, levels = c("White", "Non-White")),
         victim_age = as.numeric(as.character(victim_age))
  ) %>% 
  filter(city_state != "Dallas, TX" & city_state != "Phoenix, AZ" & city_state != "Kansas City, MO" & city_state != "Tulsa, AL")
 
  
```

The following code chunk runs a logistic regression on the dataset and extracts the OR comparing the odds of having a solved homicide for Non-Whites vs. Whites and the corresponding 95% upper and lower confidence limits in Baltimore, MD.  

```{r}
fit_logistic = 
  homicides %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial())

fit_logistic %>% 
  broom::tidy() %>% 
  janitor::clean_names() %>% 
  mutate(OR = exp(estimate),
         lower_cl = exp(estimate - (1.96 * std_error)),
         upper_cl = exp(estimate + (1.96 * std_error))
  ) %>%
  filter(term == "victim_raceNon-White") %>% 
  select(OR, lower_cl, upper_cl) %>% 
  knitr::kable(digits = 3)

```

The following code chunk creates a function to run a logistic regression and extract the OR and 95% CI. 

```{r}

or_race = function(x) {
  
fit_logistic = 
  homicides %>% 
  filter(city_state == x) %>% 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial())

fit_logistic %>% 
  broom::tidy() %>% 
  janitor::clean_names() %>% 
  mutate(OR = exp(estimate),
         lower_cl = exp(estimate - (1.96 * std_error)),
         upper_cl = exp(estimate + (1.96 * std_error))
  ) %>%
  filter(term == "victim_raceNon-White") %>% 
  select(OR, lower_cl, upper_cl) %>% 
  knitr::kable(digits = 3)
  
}

```


### Problem 2

The following code chunk reads in the data.

```{r}
birthweight = read.csv("./data/birthweight.csv") %>% 
  mutate(
    babysex = case_when(
      babysex == 1 ~ "Male",
      babysex == 2 ~ "Female"),
    babysex = fct_relevel(babysex, "Male"),
    frace = case_when(
      frace == 1 ~ "White",
      frace == 2 ~ "Black",
      frace == 3 ~ "Asian",
      frace == 4 ~ "Puerto Rican",
      frace == 8 ~ "Other",
      frace == 9 ~ "Unknown"
      ),
    frace = fct_relevel(frace, "White"),
    mrace = case_when(
      mrace == 1 ~ "White",
      mrace == 2 ~ "Black",
      mrace == 3 ~ "Asian",
      mrace == 4 ~ "Puerto Rican",
      mrace == 8 ~ "Other"
      ),
    mrace = fct_relevel(mrace, "White"),
    malform = case_when(
      malform == 1 ~ "Present",
      malform == 0 ~ "Absent"
    ),
    malform = fct_relevel(malform, "Absent")
    )
  

birthweight %>% 
  ggplot(aes(x = gaweeks, y = bwt)) +
  geom_point()

model_2 = lm(bwt ~ blength + gaweeks, data = birthweight)

model_3 = lm(bwt ~ bhead + blength + babysex + bhead*blength*babysex + bhead*blength + bhead*babysex + blength+babysex, data = birthweight)
```

